\chapter{Division of tasks}\label{c:division}

In this chapter we list individual contributions of all team members. It is important to note that a significant amount of work in this project consisted of analyzing and documenting the original implementation. This was a collaborative effort, so contributions in this area are not listed below.

\section{Individual contributions}
\subsection*{Jakub Bednarz}
\begin{itemize}
    \item Designing and implementing the \texttt{Simulation} class. 
    \item Implementing I/O and the model preparation layer.
    \item Implementing the Verlet list with the cell-based construction algorithm.
    \item Writing documentation of the program.
    \item Writing sections \ref{new:design}, \ref{new:abs}, \ref{new:docs}, \ref{summary:speed} of this thesis.
    \item Designing the structure of the repository.
\end{itemize}
\subsection*{Jakub Boguta}
\begin{itemize}
    \item Designing and implementing the parallel model for computing forces.
    \item Optimizing noise generation.
    \item Running benchmarks.
    \item Testing and patching the differences between the reference and the new program.
    \item Writing chapter \ref{c:introduction} and sections \ref{new:features}, \ref{new:parallel}, \ref{new:opt} of this thesis.
    \item Implementing the \emph{quasi-adiabatic} potential.
\end{itemize}
\subsection*{Paweł Kopaczyk}
\begin{itemize}
    \item Designing UI.
    \item Designing and prototyping additional GUI. However, this feature was ultimately discarded. 
\end{itemize}
\subsection*{Juliusz Pham}
\begin{itemize}
    \item Designing and implementing checkpoints.
    \item Writing chapter \ref{c:model} of this thesis.
    \item Implementing functions analyzing the final conformation.
\end{itemize}
\subsection*{Jan Urbanek}
\begin{itemize}
    \item Implementing the \emph{pseudo-improper dihedral} potential and other simpler potentials. 
    \item Writing chapter \ref{c:ref_impl} of this thesis.
    \item Implementing self-avoiding random walk, which generates initial conformation without self-intersections.  
\end{itemize}
